Ising Sad paper

### 2019-05-20 papers/histogram/ising.tex
working on writing the ising paper.

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

@@

\subsection{Ising Model}
% Flat Histogram Method of Wang-Landau and N-Fold Way (Schulz Binder Muller)
% https://doi.org/10.1142/S0129183102003243
\jpsays{talk about papers that covered Ising model in context of flat histogram
as an introduction to this.}
We test the convergence of several flat histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}. The $N\times N$ spin system can
take on values of $\sigma_i = \pm 1$ for up or down spins respectively. We
can write the Hamiltonian as follows:
\begin{align}
\mathcal{H} = -\sum_{<i,j>} \sigma_i \sigma_j
\end{align}
where the sum is over all enumerated spin sites. The 2D Ising spin-lattice
system is widely used as a testbed when benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition}. The
2nd order phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently
interesting for such theoretical comparisons. It is also of important note that
direct comparison of the other methods can be made with WL as its original implementation
was done on this system~\cite{wang2001determining,wang2001efficient}.

\subsection{Wang-Landau}
Wang-Landau here.

>>>>

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

@@

\subsection{Ising Model}
The 2D Ising spin-lattice system is widely used as a testbed when
benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition}. The 2nd order
phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact} make the
system sufficiently interesting for such theoretical comparisons. It is
also of important note that direct comparison of the other methods can
be made with WL as its original implementation was done on this
system~\cite{wang2001determining,wang2001efficient}.
We test the convergence of several flat histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}.
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the infinite volume limit ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over all enumerated spin sites. Beale showed that for finite
lattices the Density of States could directly be calculated from the partition
function~\cite{beale1996exact}
\begin{align}
Z = \sum_E g(E) e^{-{\beta E}}
\end{align}
where $g(E)$ is the multiplicity of the system which is identical to
$D(E)$. We can estimate the average deviation from the exact solution
using the relation~\cite{schneider2017convergence, shakirov2018convergence,
barash2017control}:
\begin{align}
\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
\end{align}
As per the implementation by Wang and
Landau~\cite{wang2001determining}, the total number of available energy
states are $N-1$.
\jpsays{Also the necessary condition that the ground state is
$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
density of states at low temperatures. If $2^N$ states is the condition
used for the ferromagnetic ising model, the rescaled factor is
determined primarily by the maximum density of states impacing the
accuracy of the energy levels at or near the ground state.}

\subsection{Wang-Landau}
The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
landau2014guide} explores energy space by initially setting $\gamma=1$,
and then decreases $\gamma$ in discrete stages.  We track the number of
moves ending at each energy and store the counts in a histogram.  For a
sufficiently flat energy histogram, $\gamma$ is decreased by a specified factor of
$\frac12$.  Flatness is defined by the ratio between the minimum
value of the histogram and its average value.  If this flatness
reaches a specified threshold (typically 0.8), the $\gamma$ value is
decreased and the histogram is reset to zero. The entire process is repeated
until $\gamma$ reaches a desired cutoff.

%The Wang-Landau approach thus has three parameters that need be
%specified: the factor by which to decrease $\gamma$ when flatness is
%acheived, the flatness criterion, and the cutoff that determines when
%the computation is complete.  In addition, an energy range (or in
%general, a \emph{set} of energies) must be supplied, so that the
%flatness criterion can be defined.

Although the algorithm is both efficient and widely used , it
suffers from a few shortcomings.  Frist, an energy range of interest must be
specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
which may require multiple simulations. Second, while Wang-Landau
converges quickly, it does not in general converge to the true density
of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
and does decrease $\gamma$ so quickly that it will never (for any
cutoff value) decrease the error in the density of states beyond a
given nonzero value.
